---
title: "Final Project Step 4"
author: "Jingru Zhang"
date: "5/26/2020"
output: word_document
---

Upload the data set and check any missing value.
```{r, message=FALSE, warning=FALSE}
library(VIM)

data <- read.csv("D:/Downloads/winequality-red.csv")
str(data)
aggr(data)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)

glimpse(data)
summary(data)
```

Explored the distribution of red wine quality and the count of each ordinal quality.
```{r, message=FALSE, warning=FALSE}
library(ggplot2)

table(data$quality)

ggplot(data, aes(x=quality)) +
  geom_bar(stat="count") +
  ggtitle("Distribution of Red Wine Quality")
```

The next step is to see the correlation between variables. There is no perfect collinearity.
```{r, message=FALSE, warning=FALSE}
library(corrplot)

cor <- cor(data)
corrplot.mixed(cor)
```

Find out the relationship between red wine quality with other indepedent variables. 
```{r, message=FALSE, warning=FALSE}

data$quality <- as.factor(data$quality)

ggplot(data, aes(x = fixed.acidity, y = quality, color = quality)) +
  geom_point()

ggplot(data, aes(x = volatile.acidity , y = quality, color = quality)) +
  geom_point()

ggplot(data, aes(x = citric.acid, y = quality, color = quality)) +
  geom_point()

ggplot(data, aes(x = chlorides, y = quality, color = quality)) +
  geom_point()

ggplot(data, aes(x = free.sulfur.dioxide, y = quality, color = quality)) +
  geom_point()

ggplot(data, aes(x = total.sulfur.dioxide, y = quality, color = quality)) +
  geom_point()

ggplot(data, aes(x = density, y = quality, color = quality)) +
  geom_point()

ggplot(data, aes(x = pH, y = quality, color = quality)) +
  geom_point()

ggplot(data, aes(x = sulphates, y = quality, color = quality)) +
  geom_point()

ggplot(data, aes(x = alcohol, y = quality, color = quality)) +
  geom_point()

```

split data
```{r, message=FALSE, warning=FALSE}
library(caTools)

data <- read.csv("D:/Downloads/winequality-red.csv")
split <- sample.split(data, SplitRatio = 0.7)
train <- subset(data, split == TRUE)
test <-  subset(data, split == FALSE)
```

Multiple linear regression
```{r, message=FALSE, warning=FALSE}
model1 <- lm(quality ~ ., data=train)

summary(model1)
```

I choose those significant independent variables only.
```{r, message=FALSE, warning=FALSE}
model2 <- lm(quality ~ volatile.acidity + chlorides+ total.sulfur.dioxide + sulphates + alcohol, data=train)

summary(model2)
```

Predict the wine quality on the testing data.
```{r, message=FALSE, warning=FALSE}
pre <- predict(model2, test, type = "response")
```

Calculate the prediction accuracy. 
```{r, message=FALSE, warning=FALSE}
actuals_preds <- data.frame(cbind(actuals=test$quality, predicteds=round(pre)))
correlation_accuracy <- cor(actuals_preds)
correlation_accuracy
```

Ordinal logistic regression
```{r, message=FALSE, warning=FALSE}
library(MASS)
library(caTools)
data$quality <- as.ordered(data$quality)
split <- sample.split(data, SplitRatio = 0.7)
train2 <- subset(data, split == TRUE)
test2 <-  subset(data, split == FALSE)
```

Use all the variable.
```{r, message=FALSE, warning=FALSE}
model3 <- polr(quality ~ .,data=train2, Hess=TRUE)
summary(model3)
```

Calculate p-value.
```{r, message=FALSE, warning=FALSE}
ctable <- coef(summary(model3))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE * 2 )
ctable <- cbind(ctable, "p value" = p)
ctable
```

I choose those significant independent variables only.
```{r, message=FALSE, warning=FALSE}
model4 <- polr(quality ~ volatile.acidity + chlorides + total.sulfur.dioxide + density + sulphates + alcohol, data=train2, Hess=TRUE)
summary(model4)
```

Calculate p-value.
```{r, message=FALSE, warning=FALSE}
ctable <- coef(summary(model4))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE * 2 )
ctable <- cbind(ctable, "p value" = p)
ctable
```

Predict the wine quality on the testing data.
```{r, message=FALSE, warning=FALSE}
pred <- predict(model4, test2)
tab <- table(pred, test2$quality)
tab
```

Calculate the prediction accuracy.
```{r, message=FALSE, warning=FALSE}
1 - sum(diag(tab))/sum(tab)
```



